\sys: Automated Data-Parallel Shell Processing

### 2020-05-02 pldi/draft.tex
Evaluation edits

% For comments
\newcommand{\eat}[1]{}
\newcommand{\TODO}[1]{\hl{\textbf{TODO:} #1}\xspace}
\newcommand{\todo}[1]{\hl{#1}\xspace}
\newcommand{\nv}[1]{[{\color{cyan}#1 --- nv}]}
\newcommand{\kk}[1]{[{\color{magenta}#1 --- kk}]}
\newcommand{\km}[1]{[{\color{blue}#1 --- km}]}
\newcommand{\review}[1]{{\color{red}#1}}
\newcommand{\tr}[1]{} %% Text and comments for technical report

@@

Micro-bench\-marks~\sx{micro} are one-off, simple one-liners that take a few seconds to write;
  such pipelines are usually composed interactively to solve a task at hand, test a hypothesis, or ``smoke out bugs''~\cite{bentley1986literate}---but, with \sys, applied to very large data sets.
These pipelines are taken from real use cases~\cite{bentley1985spelling, bentley1986literate, taylor2004wicked} and, precisely due to their small size, highlight stress on a handful of stages.

@@

\subsection{Shell One-liners}
\label{micro}

@@

\kk{Hypothesis on results. Dish gives a great speedup if one step of
  the pipeline is what takes the most time. If the pipeline is long
  with many computation heavy tasks, task based parallelism gets you
  part of the way there, so we can't expect Dish to give performance
  benefits equal to the amount of parallelism that we give it.}

Tab.~\ref{tab:eval} summarizes the collection of micro-benchmarks used to evaluate \sys.
Benchmark names (col. 1) correspond to the names on the plots (Fig.~\ref{fig:microbenchmarks}), and structure (col. 2) summarizes the different classes of commands used in the script.
% The centralized pipelines read files of various sizes from disk and write to disk.
Input sizes and times (col. 3, 4) summarize the characteristics of the largest sequential executions. % (200 files in Fig.~\ref{fig:microbenchmarks}).

Script sizes (col. 5) report on \sys's rewritten output for three different distribution sizes---$1\times$, 10$\times$, and $100\times$.
Size grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
The first number alone is interesting, as it captures \sys's output for the sequential script.
As there is no parallelism involved, it highlights the initial (bare minimum) cost corresponding to using \sys.
% this is the size of 
This change in size does not necessarily translate to observed runtime, as most of the size increase comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
FIFOs are simply pipes made explicit;
  the performance characteristics, including in-kernel buffering and synchronization mechanisms, are identical between the two.

In terms of benchmarks, \ttt{grep} is a short script centered around an expensive \sta command;
% a DFA-based regular expression matching phase;
  while the \unix \ttt{grep} command defaults to a Thompson NFA, this particular expression makes use of DFA-based backtracking patterns resulting in high runtime overhead.
The \ttt{sort} pipeline is equally short, but its central operation is a command in \pur.
The next two benchmarks, \ttt{wf} and \ttt{top-n}, are based on McIlroy's now classic word-counting program~\cite{bentley1986literate};
  they use sorting, rather than tabulation, to identify high-frequency terms in a corpus.
% \tr{The \ttt{bi-gram} pipeline calculates n-grams of base two;
%   it makes clever use of \ttt{tail} to shift a stream by one word and \ttt{paste} to fuse (zip) two streams together.}
The next pipeline is based on the original \ttt{spell} program developed by Johnson~\cite{bentley1985spelling}---another \unix classic:
  after some preprocessing, it makes clever use of \ttt{comm} to report words not in the dictionary.
Finally, \ttt{shortest-scripts} is a pipeline~\cite[pg. 7]{taylor2004wicked}that identifies the 15 shortest shell scripts in the user's \ttt{PATH};
   interestingly, it uses the \ttt{file} utility and a higher-order \ttt{wc} via \ttt{xargs}.

% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents the speedup gained by \sys as a function of the level of parallelism (1--200$\times$).
% we experiment with a $200\times$ level of parallelism to see potential overheads
Each one of these plots reports on three different configurations:
  (i) distributed, which captures only the runtime of the distributed computation,
  (ii) +compile, which adds the \sys' analysis and generation overhead, and
  (iii) +merge, which adds a final merger  gathering the results at the end of the pipeline.
Parallelism of 200$\times$ was explicitly chosen to show the \sys-internal overheads on suboptimal configurations;
  \sys itself would never attempt a factor that is above the number of available compute nodes.
The results show significant speedups, ranging between 4--70$\times$, depending on the distributability characteristics of individual pipelines.

@@

\heading{Workload}
The goal of this subsection is to evaluate \sys on an existing set of \unix pipelines found in the wild.
In a recent event celebrating \unix's 50-year legacy, Bell Labs created a set of challenges

@@

\heading{Setup}
\sys's parallelism set to \todo{16}. The input given to the pipelines
was the test input from unix50 multiplied to reach \todo{10GBs}. The
speedup of each pipeline separately is shown in
\Cref{fig:unix50-individual}. The average\todo{/median} speedup is
\todo{5.24}.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{\detokenize{./figs/unix50_individual_speedups_16.pdf}}
  \caption{ \textbf{Speedup achieved by \sys for all pipelines with
      parallelism 16. The red bars are the sequential execution time
      of each pipeline.}
    %% \todo{Scale the 0-1 region of the yaxis as much as the 1-20.}
  }
  \vspace{-15pt}
  \label{fig:unix50-individual}
\end{figure}

@@

To understand \sys better under a realistic workload, we use a large pipeline inspired by the introductory chapter of Hadoop's Definitive Guide~\cite[Chapter 2]{hadoop:15}.
At a high level, the goal is to process multi-year weather data from NCDC to identify the maximum temperature over the past several years.
The original task, as presented in the book, can be thought of comprising three sub-tasks:
  download weather data from NCDC (shell), convert them to a format that is Hadoop-friendly (shell), and process them to identify the maximum temperature (Hadoop).
Only the last of these tasks is the real focus of the chapter, demonstrating the benefits of using distributed analytics system such as Hadoop.
For \sys, we consider the full pipeline that implements all three sub-tasks---from fetching data to outputting the maximum temperature.

@@

% The first interesting result 
The effort required to write this pipeline is astonishingly low:\footnote{
  Some effort is required to understand NCDC's weather format, but this is true for any program processing this dataset.
}
  it amounts to 14 stages and, when expressed as a single line, only 240 characters.
The Hadoop program, expressed in Java and only capturing the 4 last stages, is about 137 lines of Java code---not accounting for code moving input and output in and out of HDFS.

In terms of execution time, with two Hadoop datanodes running on Hadoop 3.2.1 and Openjdk 11.0.5 without any replication (\ie fault-tolerance), the Hadoop program takes 6m22s for data processing and 1m09s for moving the dataset to HDFS.
For the same amount of parallelism, \sys takes 4m30.554s---but this includes downloading and uncompressing the data files. \kk{I think it is fine keeping the Hadoop evaluation here, since this macrobenchmark was taken from a Hadoop book.}

@@

A less obvious take-away from this case is that large, complex pipelines enable significant freedom in how to express their stages.
Several stages in the original shell program (taken from the book) are expressed as a single \ttt{awk} script;
  unfortunately, \ttt{awk} is too general to have a meaningful distributability signature.
A simpler pipeline that builds on built-in \unix primitives might be outperformed by specialized programs on a single node, but turns out to be  trivially scalable to multiple nodes.
% For example, the \ttt{sort} command at the end is an expensive version of extracting the maximum element;
%   this could have been expressed with a simple \ttt{awk} script that only keeps track of the max.

% few different ways to write some phases, 
%   for example, 
% A large pipeline resulted in a couple of 
% An interesting lesson here was that there are multiple
% First, longer pipelines can be written in multiple ways
% It is worth noting
% These are
% 
\subsection{Use Case: Wikipedia Web Indexing}
\label{macro1}

\subsection{Use Case: Docker Container Orchestration}
\label{macro1}

\subsection{Further Micro-benchmarks}

@@

>>>>

% For comments
\newcommand{\eat}[1]{}
\newcommand{\TODO}[1]{\hl{\textbf{TODO:} #1}\xspace}
\newcommand{\todo}[1]{\hl{#1}\xspace}
\newcommand{\fixme}[1]{[{\color{red}#1}]}
\newcommand{\nv}[1]{[{\color{cyan}#1 --- nv}]}
\newcommand{\kk}[1]{[{\color{magenta}#1 --- kk}]}
\newcommand{\km}[1]{[{\color{blue}#1 --- km}]}
\newcommand{\review}[1]{{\color{red}#1}}
\newcommand{\tr}[1]{} %% Text and comments for technical report

@@

Micro-bench\-marks~\sx{ours} are one-off, simple one-liners that take a few seconds to write;
  such pipelines are usually composed interactively to solve a task at hand, test a hypothesis, or ``smoke out bugs''~\cite{bentley1986literate}---but, with \sys, applied to very large data sets.
These pipelines are taken from real use cases~\cite{bentley1985spelling, bentley1986literate, taylor2004wicked} and, precisely due to their small size, highlight stress on a handful of stages.

@@

\subsection{Classic \unix One-liners}
\label{ours}

@@

\heading{Programs}
Tab.~\ref{tab:eval} summarizes the collection of micro-benchmarks used to evaluate \sys.
Benchmark names (col. 1) correspond to the names on the plots (Fig.~\ref{fig:microbenchmarks}), and structure (col. 2) summarizes the different classes of commands used in the script.
% The centralized pipelines read files of various sizes from disk and write to disk.
Input sizes and times (col. 3, 4) summarize the characteristics of the largest sequential executions. % (200 files in Fig.~\ref{fig:microbenchmarks}).
Script sizes (col. 5) report on \sys's rewritten output for three different distribution sizes---$1\times$, 10$\times$, and $100\times$.
Size grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
The first number alone is interesting, as it captures \sys's output for the sequential script.
As there is no parallelism involved, it highlights the initial (bare minimum) cost corresponding to using \sys.
% this is the size of 
This change in size does not necessarily translate to observed runtime, as most of the size increase comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
FIFOs are simply pipes made explicit;
  the performance characteristics, including in-kernel buffering and synchronization mechanisms, are identical between the two.

In more detail, \ttt{grep} is a short script centered around an expensive \sta command;
% a DFA-based regular expression matching phase;
  while the \unix \ttt{grep} command defaults to a Thompson NFA, this particular expression makes use of DFA-based backtracking patterns resulting in high runtime overhead.
The \ttt{sort} pipeline is equally short, but its central operation is a command in \pur.
The next two benchmarks, \ttt{wf} and \ttt{top-n}, are based on McIlroy's now classic word-counting program~\cite{bentley1986literate};
  they use sorting, rather than tabulation, to identify high-frequency terms in a corpus.
% \tr{The \ttt{bi-gram} pipeline calculates n-grams of base two;
%   it makes clever use of \ttt{tail} to shift a stream by one word and \ttt{paste} to fuse (zip) two streams together.}
The next pipeline is based on the original \ttt{spell} program developed by Johnson~\cite{bentley1985spelling}---another \unix classic:
  after some preprocessing, it makes clever use of \ttt{comm} to report words not in the dictionary.
Finally, \ttt{shortest-scripts} is a pipeline~\cite[pg. 7]{taylor2004wicked}that identifies the 15 shortest shell scripts in the user's \ttt{PATH};
   interestingly, it uses the \ttt{file} utility and a higher-order \ttt{wc} via \ttt{xargs}.

\heading{Performance}
% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents the speedup gained by \sys as a function of the level of parallelism (1--200$\times$).
% we experiment with a $200\times$ level of parallelism to see potential overheads
Each one of these plots reports on three different configurations:
  (i) distributed, which captures only the runtime of the distributed computation,
  (ii) +compile, which adds the \sys' analysis and generation overhead, and
  (iii) +merge, which adds a final merger  gathering the results at the end of the pipeline.
Parallelism of 200$\times$ was explicitly chosen to show the \sys-internal overheads on suboptimal configurations;
  \sys itself would never attempt a factor that is above the number of available compute nodes.
The results show significant speedups, ranging between 4--70$\times$, depending on the distributability characteristics of individual pipelines.

\heading{Discussion}
\kk{Hypothesis on results. Dish gives a great speedup if one step of
  the pipeline is what takes the most time. If the pipeline is long
  with many computation heavy tasks, task based parallelism gets you
  part of the way there, so we can't expect Dish to give performance
  benefits equal to the amount of parallelism that we give it.}

@@

The goal of this subsection is to evaluate \sys on an existing set of \unix pipelines found in the wild.

\heading{Programs}
In a recent event celebrating \unix's 50-year legacy, Bell Labs created a set of challenges

@@

the problems are well designed to highlight the \unix philosophy.
pipelines were found out in the wild and are not necessarily written in a particularly

\heading{Results}
\sys's parallelism set to \todo{16}. The input given to the pipelines
was the test input from unix50 multiplied to reach \todo{10GBs}. The
speedup of each pipeline separately is shown in
\Cref{fig:unix50-individual}. The average\todo{/median} speedup is
\todo{$5.24\times$} and the \todo{geometric mean is $XX\times$}.

\heading{Discussion}

\heading{Take-away}
\sys leads to significant performance improvements in completely unmodified \unix pipelines found out in the wild.
Small tweaks can yield further improvements, showing that \sys-awareness and scripting expertise can improve results.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{\detokenize{./figs/unix50_individual_speedups_16.pdf}}
  \caption{
	  \textbf{Speedup (left axis) and over sequential execution (right axis) for all Unix50 pipelines.}
		Parallelism is \todo{16$\times$} on \todo{10GB} of input data~\cf{unix50}.
		\nv{Can decrease height and increase width substantially?}
    %% \todo{Scale the 0-1 region of the yaxis as much as the 1-20.}
  }
  \vspace{-15pt}
  \label{fig:unix50-individual}
\end{figure}

@@

To understand \sys better under a realistic workload, we use a large pipeline inspired by the introductory chapter of Hadoop's Definitive Guide~\cite[Chapter 2]{hadoop:15}.

\heading{Program}
At a high level, the program processes multi-year weather data from NCDC to identify the maximum temperature over the past several years.
The original task, as presented in the book, can be thought of comprising three sub-tasks:
  download weather data from NCDC (shell), convert them to a format that is Hadoop-friendly (shell), and process them to identify the maximum temperature (Hadoop).
Only the last of these tasks is the real focus of the chapter, demonstrating the benefits of using distributed analytics system such as Hadoop.
For \sys, we consider the full pipeline that implements all three sub-tasks---from fetching data to outputting the maximum temperature.

@@

\heading{Results}
% The first interesting result 
The effort required to write this pipeline is astonishingly low:\footnote{
  Some effort is required to understand NCDC's weather format, but this is true for any program processing this dataset.
}
  it amounts to 14 stages and, when expressed as a single line, only 240 characters.
The Hadoop program, expressed in Java and only capturing the \emph{four} last stages, amounts to 137 lines of Java code---not accounting for commands moving data in and out of HDFS.

As this is benchmark is taken directly from the Hadoop book, it is worth reporting on its execution time.
The comparison is not unreasonable as both systems
  (i) favor message-passing rather than shared memory,
	(ii) Hadoop has no fault-tolerance for master and worker nodes (\ie both systems do not replicate data),
	(iii) the task at hand was presumably chosen by the book author as one that matches Hadoop's capabilities, and 
	(iv) the Hadoop program is written as a single program tailored for this task---with ample opportunity for optimization from the compiler and runtime system---rather than a series of loosely coupled programs that each written to handle many other cases.
\todo{The original pipeline executes in XXs}.
With two Hadoop 3.2.1 (Openjdk 11.0.5) data-nodes, the Hadoop program takes 6m22s for data processing and 1m09s for moving the dataset to HDFS.
For the same amount of parallelism, \sys takes 4m30.554s---this includes downloading and uncompressing the data files which, as described earlier, is not expressible in the Hadoop equivalent.

@@

\heading{Discussion}
Similar to Unix50~\sx{unix50}, we found that large, complex pipelines enable significant freedom in terms of expressiveness.
Several stages of the original program were expressed as a single \ttt{awk} script;
  unfortunately, \ttt{awk} is too general to have a single meaningful \todo{parallelizability} signature.
A simpler (but longer) pipeline that leverages on \unix built-ins might be outperformed by a specialized programs on a single node, but turns out to be trivially scalable to multiple nodes.
% For example, the \ttt{sort} command at the end is an expensive version of extracting the maximum element;
%   this could have been expressed with a simple \ttt{awk} script that only keeps track of the max.

\heading{Take-away}
Multi-line Java Program parallel program decomposed in two phases---a highly parallel map phase and a sequential reduce phase---corresponds to a much smaller shell script with

% few different ways to write some phases, 
%   for example, 
% A large pipeline resulted in a couple of 
% An interesting lesson here was that there are multiple
% First, longer pipelines can be written in multiple ways
% It is worth noting
% These are
% 
\subsection{Use Case: Wikipedia Web Indexing}
\label{macro2}

As a second realistic workload, we use a large pipeline for downloading and processing the web, comprised of stages collected from various repositories.

\heading{Program}
At a high level, the pipeline is comprised of two distinct components that coarsely correspond to a front- and a back-end communicating via \unix pipes.
The front-end downloads pages and extracts their URLs from their source, feeding back \todo{to} the start of the pipeline;
  the back-end applies natural-language processing---\eg trigrams, character conversion, term frequencies---to index the incoming pages.
The pipeline contains a total of 52 stages (18 for the font-end and 34 for the back-end), with written in multiple languages---for example, its \ttt{url-extract}ion utility is written in JavaScript whereas its \ttt{word-stem}ming utility is in Python.
\sys can still operate on them as their parallelizability properties---\sta for \ttt{url-extract} and \ttt{word-stem}---can be described by its DSL.

\heading{Results}
To simplify execution (and avoid hitting Wikipedia), we have saved a recent version of Wikipedia to a server within the network and to redirect all \ttt{curl} requests there.
The complexity of this pipeline makes end-to-end measurements difficult, so we measure the improvements on the three components individually and then report the differences

\heading{Discussion}

\heading{Take-away}

\subsection{Use Case: Container Orchestration}
\label{macro3}

As a third realistic workload, we use a script for orchestrating containers to test the latest built across various Linux distributions.
Compatibility testing is a common task for large software projects, commonly automated through shell scripting.

\heading{Program}
The 30-line orchestration script is unusual among our set of benchmarks in that it is not structured as a linear pipeline.
Rather, it makes extensive use of POSIX constructs common in the shell, including output redirection, conditionals, and here-documents.
It also combines commands from other classes outside \sta and \pur---more prominently \dfs---which are not parallelizable by \sys.

\heading{Results}

\heading{Take-away}

\subsection{Further Micro-benchmarks}
\label{micro}

@@

\heading{Take-away}
\sys will choose not to improve performance rather than breaking correctness; as a result, it does not always improve performance.
In highly-parallel benchmarks, it can lead to performance that is comparable to highly specialized streaming frameworks---at a fraction of the cost in terms of manual development.
